\input{../preambule.tex}


\title{Homotopy }
\author{Diego Dorn \\ Lectures notes from Christian Urech}
\date{Spring 2021}

\newcommand{\homeo}{\approx}
\newcommand{\mapspace}{\operatorname{map}}
\newcommand{\Top}{\mathrm{Top}}

\begin{document}
    \maketitle

    % \tableofcontents
    % \newpage

    \section{The compact-open topolgy}

    The \emph{aim} is to upgrade the set of (continuous) maps $f:X \to Y$
    to a space of maps, and thus the category of spaces is enriched over $\Top$ to $\Top_*$

    \begin{definition}
        The set of all maps $f: X \to Y$ is given the structure of a space map $(X, Y)$
        with the \emph{compact open topology}.
        whose subbasis is 
        \[
            B(K, U) = \bigsetst{f: X \to Y }{
                f(K) \subset U
            }
        \]
        for $K$ compact in $X$ and $U$ open in $Y$.
    \end{definition}

    Therefore a basis for this topology is given by finite intersections of $B(K_i, U_I)$, 
    $1 \leq i \leq n$, that is, 
    \[
        B(K_\bullet, U_\bullet) = \setst{f: X \to Y }{
            f(K_i) \subset U_i~
            \forall\, 1 \leq i \leq n
        }
    \]
    and an arbitrary open subset is a union of such $B(K_\bullet, U_\bullet)$.

    \begin{remark}
        We will often work with finite CW-complexes as source of map $(X, Y)$.
        If $Y$ is a CW-complex then $\mapspace(X, Y)$ has the homotopy type
        of a CW-complex, but this is not true if $X$ is an infinite CW-complex in general.
    \end{remark}

    \begin{examples}
        \begin{itemize}
            \item $X = \star$ then $\mapspace(\star, Y) \homeo Y$
            \item $X = \set{1, \dots, n}$ is discrete, then $\mapspace(X, Y) \homeo Y^n$.
            \item $X = I = [0, 1]$, then $\mapspace(I, Y)$ is the space of maps paths in $Y$.
                One can choose a nicer basis: with $t_1 = 0 < t_2 < \dots < t_n = 1$,
                a basis is $B\parenthesis{[t_i, t_{i+1}], U}$ for all $U$.
            \item $X = S^1$, then $\mapspace(X, Y) = \Lambda Y$ is the space of all free (base point is arbitrary) loops.
        \end{itemize}
    \end{examples}

    \begin{remark}
        Exponential law. We write $Y^X$ for $\mathrm{Mor}_{set}(X, Y)$.
        There is a natural bijection in $Set$
        \[
            Z^{X \times Y} \equiv {(Z^Y)}^X
        \]
        given by
        \[
            b(\p) = f: X\times Y \to Z
            <-> 
            \func{a(f) = \p}{X}{Z^Y}
                {x}{f(x, -)}
        \]
        In fact $X \times -$ is left adjoint to $Z^-$.
        Inside $Y^X SUPSET \mapspace(X, Y)$ we ask if the adjonction 
        is compactible with the compact-open topology.
        When do we have $\mapspace(X \times Y, Z) \homeo \mapspace(X, \mapspace(Y, Z))$ ?
        This isn't always the cas, so we will impose restrictions.
    \end{remark}

    \begin{proposition}~
        \begin{itemize}
            \item Let $X$ be locally compact, then the evaluation $ev$ is continuous:
            \[
                \func{ev}{\mapspace(X, Y) \times X}{Y} {(f, x)}{f(x)}
            \]
            \item If $Y$ is locally compact, a map $F: X \times Y \to Z$ is continuous
                iff $a(f): X \to \mapspace(Y, Z)$ is continuous.
        \end{itemize}
    \end{proposition}

    \begin{proof} We prove both parts:
        \begin{itemize}
            \item 
                Let $U \ni f(x)$ be open in $Y$. Since $f$ is continuous,
                $f^{-1}(U)$ is an open neighbourhood of $x$. Since $X$ is locally compact,
                there is $x \in K \subset f^{-1}(U)$, with $K$ compact.
                We choose $B(K, U)$ in $\mapspace(X, Y)$ and $K$ as neighbourhood of $x \in X$.
                \[
                    ev(B(K, U) \times K) \subset U
                \]
                so $ev$ is indeed continuous.
            \item Notice first the if $f: X \times Y \to Z$ is a map,
                then \todo{fill}
                Next we show that $a(f)$ is continuous by looking at the preimage of 
                $G(L, W)$. For $L \subset Y$ compact, $W$ open in $Z$, 
                \[
                    a(f)^{-1}(B(L, W)) = \setst{x \in X}{f(x \times L) \subset W}.
                \]
                Since $W$ is open, $f^{-1}(W)$ is open in $X \times Y$ and contains 
                $x \times L$.
                Let us choose an open subset $U \times V$ for the product topology
                in $X \times Y$. Since $L$ is compact, we ccan assume that is $x \times L$.
                We have shown that every point $x \in a(f)^{-1}(B(L, W))$
                admits an open neighbourhood in this preimage.
            
                Last part: Assume $\p: X \to \mapspace(Y, Z)$ is continuous. 
                We prove that $b(\p)$ is continuous.
                \[
                    b(\p): 
                        X \times Y
                        \xto{\p \times 1}
                        \mapspace(Y, Z) \times Y
                        \xto{ev} Z
                \]
                \[
                    (X, y) \mapsto (\p(x), y) 
                    \mapsto \p(x)(y)
                \]
                So $b(\p)$ is continuous because $\p \times 1$ is so
                amd $ev$ is also continuous by the first point.
        \end{itemize}
    \end{proof}

    \begin{lemma}
        Let $X, Y$ be Hausdorff spaces. Then a subbasis for $\mapspace(X \times Y, Z)$
        is given by $B(K \times L, W)$
        where $K$ is compact in $X$, $L$ is compact in $Y$ and $W$ is open in $Z$.
    \end{lemma}

    \begin{proof}
        Let $A$ be compact in the product, $A_X = \pi_X(A)$,
        $A_Y = \pi_Y(A)$ are compact (projections of a compact).
        We have $A \subset A_X \times A_Y$.
        For $f \in B(A, W)$, in principle $f \notin B(A_X, A_Y, W)$
        For any $(x, y) \in A$, $(x, y) \in A_X \times A_Y \cap f^{-1}(W)$.
        Choose $(x, y) \in U_x \times U_y$ with
        $x \in U_x \in A_X$ open
        and
        $y \in U_y \in A_Y$ open.

        Now, compact Hausdorff implies normal, hence regular so 
        we can separate $x$ form $A_X \setminus U_x$ by $x \in U \subset \bar{U} \subset U_x$.
        the same goes for $y$ form $A_Y \setminus U_y$ by $y \in V \subset \bar{V} \subset U_y$.

        Finally, the compact subspaces $\bar{U} \times \bar{V}$ cover $A$, so we can extract
        a finite cover $
            \bar{U}_i \times \bar{V}_i
        $, $1 \leq i \leq n$.
        Then $f \in \bar{U}_i \times \bar{V}_i, W)$
        since $ \bar{U}_i \times \bar{V}_i \subset A$, so
        $f \in \bigcap_{i = 1}^n \bar{U}_i \times \bar{V}_i$
    \end{proof}

    Notice that $\a$ sends $B(K \times L, W)$ to $B(K, B(L, W))$.

    \begin{lemma}
        If $Z$ is Hausdorff, then $\mapspace(Y, Z)$ is Hausdorff.
    \end{lemma}

    \begin{lemma}
        If $Z$ is Hausdorff, $\mathcal{B}$ is a subbasis of its topolgy,
        then $B(K, \mathcal{B})$ for $K \subset Y$ compact form a 
        subbasis for the compact-open topology on $\mapspace(Y, Z)$.
    \end{lemma}



    \newpage
    .
\end{document}
