\input{../preambule.tex}

\title{Homework 2}
\author{Diego Dorn}
% \date{Automn 2021}

\newcommand{\klotski}{\textsc{Klotski}\xspace}
\newcommand{\UniqueSAT}{\textsc{UniqueSat}\xspace}
\newcommand{\UnSAT}{\textsc{UnSat}\xspace}
\newcommand{\SAT}{\textsc{Sat}\xspace}

\begin{document}
    \maketitle

    \section*{Exercise 1}
    Using Savtich's theorem, it suffices to show that 
    $\klotski \in \NPSPACE$. 

    An non deterministic algorithm that solves \klotski
    in polynomial space is the following (in Python notation):

    \begin{lstlisting}[language=python]
        def klotski(board):
            for i in range(2 ** board.size):
                if red_piece_on_exit(board): 
                    return True
            
                guess next_move
                apply(next_move, board)
            return False
    \end{lstlisting}

    Where \texttt{guess} chooses a move randomly and \texttt{apply}
    modifies the board according to the chosen move. It is clear that
    \texttt{apply} can be implemented to use a space polynomial
    in the size of the board configuration and similarly for 
    \texttt{red\_piece\_on\_exit}.
    Thus, this algorithm operate with polynomial space.

    This algorithm always terminates with the right answer since
    if a board can be solved, it can be solved without going twice 
    through the same configuration, and thus all boards can be solved 
    in less than $2^n$ moves, where $n$ is the length of the input, 
    assuming that for all board configuration on a given square
    their description have the same length $n$.

    \section*{Exercice 2}

    We first show that \UniqueSAT is $\coNP$-hard.
    To that extent, we reduce \UnSAT to it.

    Let $\p$ be a DNF formula which is an input of \UnSAT,
    we construct a $\p'$ that has a unique solution if and only if 
    $\p$ is unsatisfiable. 
    Let $(C_i)_{i \leq m}$ be the $m$ clauses of $\p$,
    $x_1, \dots, x_n$ be the variables
    appearing in $\p$ and $z$ be a new variable.
    Now we set \[
        \p' = 
            \bigwedge_{i = 0}^n (\bar z \vee x_i)
            \wedge
            \bigwedge_{i = 0}^m (z \vee C_i)
    \]

    The assignements that satisfy this formulas are:
    \begin{itemize}
        \item if $z \mapsto 1$, the only solution is $x_i \mapsto 1$ for all $i$
            because of the first part of the formula.
        \item if $z \mapsto 0$, the solutions are the same as the solutions for $\p$
            with the added assignement of $z$.
    \end{itemize}

    Thus, if $\p$ is unsatisfiable then $\p'$ has exactly one assignement, with $z \mapsto 1$
    and if $\p$ has at least one satisfying assignement, it has exactly one more.
    Therefore $\p \in \UnSAT \iff \p' \in \UniqueSAT$
    and $\UnSAT \leq \UniqueSAT$.

    We have the following chain of reductions \[
        \SAT \geq \UniqueSAT \geq \UnSAT \geq \compl\UniqueSAT \geq \SAT
    \]
    where the first reduction is the hypothesis of the exercice,
    we just proved the second, the third comes from the fact that 
    if $\UniqueSAT \in \NP$, then $\compl\UniqueSAT \in \coNP$
    and the last reduction is the same as the second, 
    but with complements of the languages.

    All together, we have an equivalence between $\SAT$ and $\UnSAT$
    and $\NP = \coNP$.


    \section*{Exercice 3}

    Let $f : \set{0, 1}^n \to \Z$ be any function, $t < 2^n$ an integer
    and $s = 2^{n} / t$ represents the maximum number of steps of the algorithm.
    We use the following algorithm to find a local minimum of $f$:
    \begin{itemize}
        \item let $x_1 = \mathrm{argmin}_{i < t} f(r_i)$
            where $r_i \in \set{0, 1}^n$ are random strings.
        \item for $i = 1, \dots, s$
        \begin{itemize}
            \item let $x_{i+1} = \mathrm{argmin}_{y} f(y)$, where the $\mathrm{argmin}$
                is taken over the $n$ strings that differ from $x$ by one bit.
            \item if $f(x_i) \leq f(x_{i+1})$, we have found a local minimum 
                so we stop and return $x_i$ (no need to make any query at this step, 
                we can store the results of $f$ we already queried before).
            \item otherwise continue the loop
        \end{itemize}
        If we did not find a local minimum in the $t$ steps, we have failed
        so ew return a random string.
    \end{itemize}

    When are we sure that this randomised algorithm produces a correct result ?
    Let $p_0, \dots, p_{2^n}$ be an enumeration of the strings in $\set{0, 1}^n$
    such that for all $i <j$ we have $f(p_i) \leq f(p_j)$.
    Now the string $x_0$ is one of the $p_i$, so let $N$ be such that $x_0 = p_N$.
    
    We show that if $N < s$, then the algorithm will produce the correct value.
    Indeed at each step, $f(x_{i+1}) < f(x_i)$, so we take a $p_i$
    with a smaller index everytime
    (except if the algorithm stops early, but then there we are done).
    Thus we either reach $p_0$, a global minimum in less than $s$ steps
    or a local minimum earlier.

    Now let $S = \set{p_0, \dots, p_{t-1}}$.
    The failure probability is 
    \begin{align*}
        \IP(\forall i < t, r_i \notin S) 
        &= \prod_{i < t} \IP(r_i \notin S) & (\text{independance})
        \\&= \IP(r_0 \notin S)^t              & (\text{same law})
        \\&= \parenthesis{1 - \frac{\abs{S}}{2^n}}^t
        \\&= \parenthesis{1 - \frac{2^{n}}{t2^n}}^t
        \\&= \parenthesis{1 - \frac{1}{t}}^t
        \\&< e^{-1} < 1/2
    \end{align*}

    Where the last line follows from the fact the 
    sequence $\parenthesis{1 - 1/n}^n$ converges to $e^{-1}$ from below.
    Therefore, our algorithm is a randomised bounded error algorithm.

    Finally, we see that the algorithm makes $t$ queries at the start
    then at most $n$ for each of the (at most) $s$ iterations of the loop,
    so the total number of queries $T \leq sn + t = ns^n/t + t$.
    In order to minimize $T$ we can set $t = 2^{n/2}\sqrt{n}$
    and we have that $T = \O{2^{n/2}\sqrt{n}}$ as required.


    \newpage
    .
\end{document}
