\input{../preambule.tex}

\usepackage[backend=biber,style=alphabetic]{biblatex}
\addbibresource{biblio.bib}

\title{Homework 1}
\author{Diego Dorn}
% \date{Automn 2021}

\begin{document}
    \maketitle

    % \tableofcontents

    \section*{Exercise 1}

    For a decisive Turing machine $\Mm$, we write $L(\Mm)$ for the language it decides.
    We first show that if $L(\Mm) \in \NP$, then $L(\Mm) \in \coNP$.

    \begin{claim}
        If 
        $L(\Mm) \in \NP$ 
        for all decisive Turing machine $\Mm$, 
        then for all decisive Turing machine $\Nn$,
        $L(\Nn) \in \coNP$.
    \end{claim}

    \begin{proof}
        Let $\Mm$ be a decisive Turing machine,
        we build a decisive Turing machine $\compl{\Mm}$
        that runs exactly the same way as $\Mm$, but on each branch,
        when it halts, it outputs \textit{yes} when $\Mm$ outputs \textit{no}
        and \textit{no} when $\Mm$ outputs \textit{yes}. The \textit{maybe} 
        do not change. This way, $\compl{\Mm}$ accepts a word $x$ if and only
        if $\Mm$ does not accept $x$. 
        Thus $L(\compl{\Mm}) = \compl{L(\Mm)}$.
        Using the hypothesis with $\compl{\Mm}$, 
        we get that $\compl{L(\Mm)} = L(\compl{\Mm})  \in \NP$
        and therefore $L(\Mm) \in \coNP$.
    \end{proof}

    We now prove that for a polytime decisive Turing machine $\Mm$, 
    $L(\Mm) \in \NP$.
    To that extent, we construct a polytime verifier $V(x, C)$ for $L(\Mm)$
    as follows:
    \begin{itemize}
        \item If $C$ doesn't encode a sequence of transitions of $\Mm$, reject $x$. 
            Otherwise, let $(C_n)$ be the finite sequence of transitions encoded by $C$.
        \item Simulate $\Mm$ on input $x$ and on the $n$-th step, use $C_n$ to decide 
            of the next state of the deterministic simulation. 
            If the transition $C_n$ is not in the transition function of $\Mm$, reject $x$.
        \item When the simulation of $\Mm$ halts, if it outputs $yes$, 
            accept $x$. Otherwise, reject $x$.
    \end{itemize}

    Since we can simulate Turing machines with only a polynomial slowdown,
    and $\Mm$ is a polytime NTM, 
    it is clear that $V$ run in polynomial time.
    Thus, we only need to prove that $\Mm$ accepts $x$ if and only if there exist $C$
    such that $V(x, C)$ accepts $x$. 

    First, we notice that on any given input $x$, there cannot be one branch 
    of the computation that outputs \textit{yes} and an other one that outputs \textit{no}.
    
    For the first direction, there is a branch of the computation of 
    $\Mm$ on input $x$ that halts on a \textit{yes},
    so it suffices to take $C$ as the sequence of transitions
    that make up that branch.

    For the reverse direction, if there is a $C$ that make $V$ output \textit{yes},
    it must be a valid sequence of transitions for $\Mm$, 
    and when $\Mm$ follows this branch, it outputs \textit{yes}.
    Since there is a branch of $\Mm$ that outputs \textit{yes},
    all the other branches must outputs \textit{yes} or \textit{maybe}
    and thus $\Mm$ accepts $x$.

    Therefore, $L(\Mm) \in \NP$ and by the lemma, we have that for all 
    decisive Turing machine $\Mm$, $L(\Mm) \in \NP \cap \coNP$.


    \section*{Exercice 2}

    First we show that \textsc{SepMatch} is in $\NP$.
    We can define a verifier $V(G, k, C)$ 
    for \textsc{SepMatch} with the certificate $C$ being
    an encoding of the edges of the matching.
    $V$ checks for all pair of edges if the distance
    between each pair of points is less than 2 in $G$, 
    and that there are exactly $k$ edges. This can be done
    in $\O{|G|^3}$.

    To show that \textsc{SepMatch} is $\NP$-hard,
    we reduce \textsc{IndSet} to it.

    Let $G = (V, E)$ be any finite graph. We define the graph $G' = (V', E')$
    as \begin{itemize}
        \item $V' = V \times \set{0, 1}$
        \item $E' = \setst{
            \bigparenthesis{(u, i), (v, j)} \in V' \times V'
        }{
            (u, v) \in V \vee (u = v \wedge j \neq i)
        }$, that is, two vertices are connected in $G'$
        if their first coordinates are connected in $G$
        or if they are different but have the same first coordinate.
    \end{itemize}

    Our reduction from \textsc{IndSet} to \textsc{SepMatch}
    if the function that maps an input $\angles{G,k}$ of \textsc{IndSet}
    to the input $\angles{G',k}$ of \textsc{SepMatch} .

    Before proving that this is indeed a reduction, we notice that this map 
    is computable in polynomial time and we setup some notations and a lemma.
    Let $d_G$ and $d_{G'}$ the distances between vertices of $G$ and $G'$
    respectively. For a node $u \in G'$, we write $u = (u_0, u_1)$,
    and $\pi_0: G' \to G$ the projection on the first coordinate, so $\pi_0(u) = u_0$.

    \begin{lemma*}
        For all $u, v \in V'$ with $u_0 \neq v_0$, \[
            d_{G'}(u, v) = d_G(u_0, v_0).
        \]
    \end{lemma*}
    \begin{proof}
        Let $u, v \in V'$. We first show that $d_{G'}(u, v) \geq d_G(u_0, v_0)$.
        Indeed, let $\w = \w_0\w_1\dots\w_n$ be a path from $u$ to $v$ in $G'$. 
        We have that $\w_G = \pi_0(\w_0) \pi_0(\w_1) \dots \pi_0(\w_n)$
        is a path in $G$ if we remove vertices that may appear more than once consecutively.
        Thus $\abs{\w_G} \leq \abs{\w}$ and the distances in $G$ are always 
        less than the distances in $G'$.

        It remains to be shown that $d_{G'}(u, v) \leq d_G(u_0, v_0)$.
        Let $\w = \w_0\w_1\dots\w_n$  be a path from $u_0$ to $v_0$ in $G$. We
        construct a path $\w_{G'}$ from $u$ to $v$ of the same length 
        in $G'$. For $i = 0, \dots, n$,
        \[ 
            (\w_{G'})_i = \begin{cases}
                (\w_0, u_1) & \tif i = 0 \\
                (\w_i, v_1) & \tif i > 0
            \end{cases}
        \]
        If $u_0 \neq v_0$, this is a path in $G'$ from $(\w_0, u_1) = u$ to $(\w_n, v_1) = v$
        since at any step $\w_i$ is a neighbour of $\w_{i+1}$.
        Thus distances in $G'$ are at most distances in $G$.
    \end{proof}

    We now show that $\angles{G, k} \in \textsc{IndSet}$ if and only if 
    $\angles{G', k} \in \textsc{SepMatch}$.
    
    \paragraph{$\implies$}
    Let $S \subset V$ be an independent set of size $k$,
    then $M = \setst{
        \bigparenthesis{
            (s, 0), (s, 1)
        }}{
            s \in S
        }
    $ is a separated matching (of size $k$) in $G'$.
    Indeed, let $e, e' \in M$ be two different edges.
    We know that for all $u \in e$ and $v \in e'$, 
    \begin{itemize}
        \item $u_0 \neq v_0$ since the edges are different.
        \item $d_G(u_0, v_0) \geq 2$ since $S$ is an independent set.
    \end{itemize}
    So by the lemma, $d_{G'}(u, v) = d_G(u_0, v_0) \geq 2$
    and $M$ is indeed a separated matching.

    \paragraph{$\impliedby$}
    If $\angles{G', k} \in \textsc{SepMatch}$, 
    let $M$ be a separated matching of size $k$. 
    We show that 
    \[
        S = \setst{u_0}{(u, v) \in M}
    \]  
    is an independent set (of size $k$) in $G$.
    To that extent, let $u_0, u_0' \in S$
    and $(u, v), (u', v') \in M$ the corresponding edges.
    Since $M$ is a separated matching, 
    we have that $d_{G'}(u, u') \geq 2$
    so that, by the lemma, $d_G(u_0, u_0') = d_{G'}(u, u') \geq 2$,
    so $u_0$ and $u_0'$ are indeed separated.


    \section*{Exercice 3}

    \paragraph{1.}
    Let \Uu be a universal Turing machine that can
    simulate an other Turing machine with a polynomial slowdown.
    We set $A$ as the same oracle as in class, that is, 
    \[
        A = \setst{ \angles{M, 1^k} }{
            \Uu\parenthesis{
                \angles{M, \angles{M, 1^k}}
            } = 0
            \text{ in } 2^k \text{ steps}
        }.
    \]
    We have seen in class that $\Ptime^A = \NP^A$, so we just show that
    $\coNP^A = \NP^A$. Indeed, 
    \begin{align*}
        L \in \coNP^A 
        &\iff 
        \compl{L} \in \NP^A
        \\& \iff 
        \compl{L} \in \Ptime^A
        \\& \iff 
        L \in \coP^A = \NP^A.
    \end{align*}
    Were the last line follows from the fact that if a language can be 
    decided deterministically in polynomial time, then its complement
    can be too, by just inverting the outputs of the Turing machine.

    \paragraph{2.}
    Let $ \Bb = \left\{ B : \forall n \abs{B \cap \set{0, 1}^n} = 1 \right\} $,
    and for each $B \in \Bb$, we set \[
        L_B = \setst{1^n}{ \exists x \in \set{0, 1}^{n-1}, 0x \in B}.
    \]

    \begin{claim}
        For all $B \in \Bb$, $L_B \in \NP$.
    \end{claim}
    \begin{proof}
        If $1^n \in L_B$, a certificate is the $x \in \set{0, 1}^{n-1}$
        such that $0x \in B$. Indeed, a verifier can check if $0x \in B$ in a 
        single query to the oracle.
    \end{proof}


    \begin{claim}
        For all $B \in \Bb$, $L_B \in \coNP$.
    \end{claim}
    \begin{proof}
        We have that \begin{align*}
            \compl{L_B} &=
            \setst{1^n}{ 
                \forall x \in \set{0, 1}^{n-1}, 0x \notin B
            }
            \\&=
            \setst{1^n}{ 
                \exists x \in \set{0, 1}^{n-1}, 1x \in B
            }
        \end{align*}
        where the last equality follows from the the fact that
        $\abs{B \cap \set{0, 1}^n} = 1$.
        With $\compl{L_B}$ written as such, we can apply the same proof as the previous
        claim, mutatis mutandis, $\compl{L_B} \in \NP$.
    \end{proof}

    \begin{claim}
        There exist $B \in \Bb$ such that $L_B \notin \Ptime^B$.
    \end{claim}
    \begin{proof}
        Fix an enumeration $(\Mm_i, p_i)_{i \in \N}$ of all pairs
        of oracle compatible Turing machines,
        and positive polynomials\footnote{
            The trick of bounding machines via polynomials
            is due to \cite{baker1975relativizations}
            and allows a formal definition of the oracle.
            Appart from this idea, the rest of the proof is my own work.
            The first version of my proof assumed that 
            we can ignore the queries outside the range where 
            we tried to fool $\Mm_i$, but it is absolutely 
            not the case. Searching in my notes then on the
            web lead to many incorrect proofs then this one
            that I now cannot unsee
            and acknowledges.
        }.

        We would like an oracle $\bar B$
        that fools each polytime machine at least once.
        For that, we now construct iteratively a set $B$ and its complement $B'$
        that is, at each step we can add inputs to either $B$ or $B'$
        consistently and the final oracle will satisfy $\bar B \supset B$ and 
        $\compl{\bar B} \supset B'$.
        Let $n_0 = 0$.

        At step $i$, let $n > n_{i}$ such that
        $2^{n-1} > p_i(n)$.
        We run $\Mm_i$ on $1^n$ for $p_i(n)$ steps.
        At any step, if the machine asks the oracle for the string $q$:
        \begin{itemize}
            \item if $q \in B$, reply \textit{yes}.
            \item if $q \in B'$, reply \textit{no}.
            \item if $\abs{\set{x \in B' : |x| = |q|}} = 2^{\abs{q}} - 1$,
                that is, all the queries of size $|q|$ but the current one
                have been made, reply \textit{yes} and add $q$ to $B$.
            \item otherwise, reply \textit{no} and add $q$ to $B'$.
        \end{itemize}
        
        When the machine halts or after $p_i(n)$ steps,
        it has not queried all of 
        the strings of size $n$ that start with
        a 0 nor those that start with a 1,
        since it ran in $p_i(n) < 2^{n-1}$ steps.
        All its queries of size $n$ are in $B'$.
        Now, if the machine: \begin{itemize}
            \item halts on \textit{yes},
                add one string of size $n$ starting
                with 1 to $B$, and all the other strings of size $n$ to $B'$
            \item halts on \textit{no}, do the same but taking a string
                that starts with 0.
            \item does not halt in less than $p_i(n)$ steps,
                we do the same as if it halted on \textit{no}.
        \end{itemize}
        Finally, set $n_{i+1} = 2^n$ and proceed to step $i+1$. 
        Notice that neither $B$ nor $B'$
        contain strings of length $2^n$ because the machine ran for less than $2^n$
        steps.

        \paragraph{} At the end, set $\bar B$ to any set in $\Bb$ such
        that $B \subset \bar B$ and $B' \cap \bar B = \emptyset$.
        Now take any turing machine $M$ with oracle $\bar B$ 
        with runtime bounded by some polynomial $p$. 
        We have that $(M, p)$ belongs to the enumeration $(\Mm_i, p_i)_{i \in \N}$
        at, say, index $i$. 
        Now if we look at the construction of $\bar B$ at step $i$, 
        we have that the machine was ran for $p_i(n)$ steps,
        so it stoped ``naturally'' and thus its output does not
        correspond to wether $1^{n_i} \in L(\bar b)$ and $L(\bar B) \notin \Ptime^{\bar B}$.
    \end{proof}

    Using the three claims, $L_B \in \NP^B \cap \coNP^B \setminus \Ptime^B$.

    \printbibliography

    \newpage
    .
\end{document}
