\input{../preambule.tex}

\title{Homework 3}
\author{Diego Dorn}
% \date{Automn 2021}

\newcommand{\Eq}{\textsc{Eq}\xspace}
\newcommand{\Iso}{\textsc{Iso}\xspace}
\newcommand{\REF}{\textsc{Ref}\xspace}
\newcommand{\SI}{\textsc{SI}\xspace}
\newcommand{\Rcc}[1]{\mathrm{R}^{cc}_{1/3}\parenthesis{#1}}
\newcommand{\Dcc}[1]{\mathrm{D}^{cc}\parenthesis{#1}}


\begin{document}
    \maketitle

    \section*{Exercise 1}

    \begin{claim}
        $f$ has sensitivity $s(f) \leq \O{n}$.
    \end{claim}
    \begin{proof}
        We first show that $s(f) \leq \O{n}$.
        Let $x \in \set{0, 1}^{n^2}$ be an input of the problem.
        We want to show that $s(f, x) \leq \O{n}$. 

        If $f(x) = 0$,
        we show that each row has at most 3 sensitive bits, which yields the result.
        Let $w$ be the number of ways to make the row paired.
        \begin{itemize}
            \item if a row has 0, 2, or 4+ bits set, changing a bit produces 
                string of odd hamming weight and thus not paired, so $w = 0$.
            \item if a row has one bit set, flipping one of its neighbours
                are the only way to change the output, so $w \leq 2$.
            \item if a row has three bits set, flipping any of those three bits
                are the only way to get to exactly two bits set, and thus there are
                at most three ways to pair the row, so $w \leq 3$.
        \end{itemize}

        Now if $f(x) = 1$, and there is more than one row which is paired,
        then there is no way to make them all of them unpaired by flipping only one 
        bit, so $s(f, x) = 0$.

        Finally, if $f(x) = 1$, and there is exactly one paired row, the only
        ways to make the row unpaired is by flipping one of the bits of the row,
        so $s(f, x) = n$.
        
        In each case, $s(f, x) \leq 3n$ so $s(f) \leq \O{n}$.
    \end{proof}

    \begin{claim}
        $f$ has block sensitivity $bs(f) \geq \Omega(n^2)$.
    \end{claim}
    \begin{proof}
        It suffices to exibit an input $x$ of $f$ which has block-sensitivity
        $bs(f, x) = \Omega(n^2)$.
        We take $x = 0^{n^2}$, and for each $i = 1, \dots, n^2/2$
        the subsets $S_i = \set{2i-1, 2i}$.
        Clearly, the $S_i$ are disjoints and for each $i$, $f(x^{S_i}) = 1 \neq f(x) = 0$.
        Thus, $bs(f, x) \geq n^2/2$ and therefore $bs(f) \geq n^2/2 \geq \Omega(n^2)$.
    \end{proof}

    \section*{Exercice 2}
    \paragraph{a)}
    We know that $\Rcc{\Eq_{n^2}} = \O{1}$, so
    we reduce $\Iso_n$ to it, which will prove that $\Rcc{\Iso_n} = \O{1}$.
    To that extent, both Alice and Bob compute a cannonical form of their
    graph, for instance the graph in the equivalence class of their graph
    with the lexicographically least adjency matrix. This adjency matrix
    can be represented as a $n^2$ bit string, 
    and the two strings will be equal if and only if the graphs were isomorphic.

    Since the best randomized protocol for equality satisfies 
    $\Rcc{\Eq_{n^2}} = \O{1}$, the same holds for $\Iso_n$.

    \paragraph{b)}
    We know that 
    $\Dcc{\Eq_n} = \Omega(n)$
    and thus
    $\Dcc{\Eq_{n^2}} = \Omega(n^2)$.
    We reduce $\Eq_{n^2}$ to $\Iso_{7n}$
    which shows that $\Dcc{\Iso_n} = \Omega(n^2)$.

    From a $n^2$ bit string $x$, we construct the graph $G_x$
    on $7n$ vertices with labels $k_1, \dots, k_{3n}$, $s_1, \dots, s_{2n}$ and $a_1, \dots, a_{2n}$
    by describing which edges belong to the graph:
    \begin{itemize}
        \item Each pair $k_i$, $k_j$ is connected, 
            making a copy of $K_{3n}$, the complete graph on $3n$ vertices.
        \item For $i = 1, \dots, 2n$, there is an edge between 
            $a_i$ and $k_1, \dots, k_i$, 
            making $a_i$ have exactly $i$ neighbours in the copy of $K_{3n}$.
        \item For $i = 1, \dots, 2n$, there is an edge between 
            $a_i$ and $s_i$.
        \item Let $b: [n^2] \to \setst{\set{a_i, a_j}}{i \neq j \in [2n]}$ be any injection from $[n^2]$ to the set of edges 
            between $a_i$'s. This injection exists, since the size of the set of edges 
            is $\frac{2n(2n - 1)}{2} \geq n^2$.
            Finally, for each $i \in [n^2]$ such that $x_i = 1$,
            there is the edge $b(i)$.
    \end{itemize}

    For instance for $n = 3$, we get this graph:

    \begin{center}
    \begin{tikzpicture}
        \node[regular polygon, draw] (K3n) {$K_{3n}$};
        \foreach \x in {1, ..., 6} {
            \node[draw, circle] (a\x) at ($(K3n) + (\x - 3.5, -2)$) {$a_{\x}$};
            \foreach \i in {1, ..., \x} \draw (a\x) edge[out=90, in=-90+\x*20-4*20+8*\i] (K3n) ;
            \node[draw, circle] (s\x) at ($(a\x) + (0, -1.5)$) {$s_{\x}$} edge (a\x);
            \foreach \i in {1, ..., \x} 
                \draw (s\x) edge[dashed,out=-90,in=-90] (s\i);
        };
    \end{tikzpicture}
    \end{center}
    Where the edges between the $s_i$ are set or not depending on
    the string $x$. As a sanity check, note that here there are 
    15 edges that can be set which leaves enough place to encode
    the $n^2 = 3^2 = 9$ bits in the string $x$.

    We now show that the function that maps $(x, y) \mapsto (G_x, G_y)$
    is indeed a reduction from $\Eq_{n^2}$ to $\Iso_{7n}$.
    First, if $x = y$, $G_x = G_y$ by construction and the two
    graphs are isomorphic.
    
    For the reverse direction, assume that $G_x \simeq G_y$,
    and the the isomorphism is realised by $f: G_x \to G_y$.
    The two graphs are on the same set of vertices, $k_i$, $a_i$, $s_i$ 
    so $f$ corresponds to a permutation of those vertices.
    If we show that $f(s_i) = s_i$
    we will be done, as for all $i \in [n^2]$, 
    \[
        x_i = 1 \iff b(i) \in E(G_x) \iff f(b(i)) \in E(G_y)
        \iff b(i) \in G(y) \iff y_i = 1
    \],
    where the second last equivalence is because $f$ fixes each $s_i$.

    To show that $f(s_i) = s_i$ for all $i \in [2n]$, we use
    the fact that the set of vertices of $G_x$ that verify a given property
    must be mapped by $f$ to the set which verify the same
    property in $G_y$.
    We then proceed in three steps:
    \begin{itemize}
        \item Let $K = \setst{k_i}{i \in [3n]}$ be the set of all $k_i$.
            Then $f(K) = K$, since $K$ is the unique $3n$-clique in both $G_x$
            and $G_y$.
        \item For $i \in [2n]$, $f(a_i) = a_i$, indeed, for both graphs,
            $a_i$ is the only vertex adjacent to $K$ with degree $i + 1$.
        \item For $i \in [2n]$, $f(s_i) = s_i$, indeed $s_i$ is the only
            neighbour of $a_i$ not in $K$, and since both $a_i$ and $K$
            are fixed, we must have $f(s_i) = s_i$.
    \end{itemize}

    Therefore, $x = y \iff G_x \simeq G_y$ and $\Eq_{n^2}$ reduces to $\Iso_{7n}$,
    so $\Iso_{7n} \geq \Omega(n^2)$ and $\Iso_n \geq \Omega(n^2)$.\qed
    

    \section*{Exercice 3}
    We know that $\Dcc{\SI_n} = \Omega(n)$, where $\SI$ is
    the set-intersection problem,
    and therefore $\Dcc{\compl\SI_n} = \Omega(n)$,
    where $\compl\SI$ is the complement of $\SI$,
    i.e. set-non-intersection.
    We reduce $\compl\SI_n$ to $\REF_{n+1}$ to prove the lower bound.

    Let $a \in \set{0, 1}^n$ and $b \in \set{0, 1}^n$ be Alice and Bob's input
    for $\compl\SI_n$ respectively.
    We define $A = \setst{i \in [n]}{a_i = 1} \cup \set{n + 1}$
    and similarly $B = \setst{i \in [n]}{b_i = 0} \cup \set{n + 1}$.
    Notice that:
    \begin{align*}
        (a, b) \in \compl\SI_n 
        &\iff a \cap b = \emptyset 
        \\& \iff \forall i \in [n],~ a_i = 1 \implies b_i = 0
        \\& \iff A \subset B
    \end{align*}

    Our reduction to $\REF_{n+1}$ is then 
    \[
        \mathcal P = \set{A} \cup 
        \bigcup_{\substack{i \in [n] \\ a_i = 0}}
            \set{\set{i}}
    \]
    \[
        \mathcal Q = \set{B} \cup 
        \bigcup_{\substack{i \in [n] \\ b_i = 1}}
            \set{\set{i}}
    \]
    which are the partitions containing $A$ (resp. $B$) and the rest as singletons.

    It remains to show that $\mathcal P \sqsubseteq \mathcal Q \iff a \cap b = \emptyset$, but
    \begin{align*}
        \mathcal P \sqsubseteq \mathcal Q 
        & \iff \forall P \in \mathcal P, \exists Q \in \mathcal Q,~ P \subset Q
        \\& \iff \exists Q \in \mathcal Q,~ A \subset Q
        \\& \iff A \subset B
        \\& \iff (a, b) \in \compl\SI_n
    \end{align*}
    Where the second line comes from the fact that the condition is always met
    for singletons, and the only non-singleton element of $\mathcal P$
    is $A$. The third line comes from the fact that $n + 1 \in A$
    and the only set that contains $n + 1$ in $\mathcal Q$ is $B$, 
    so if $A$ is a subset of some part of $\mathcal Q$, it must be $B$.

    Thus, solving $\compl\SI_n$ is simpler than solving $\REF_{n+1}$ and 
    can be done only in $\Omega(n)$, so the same is true for $\REF_{n+1}$

\end{document}
